A pipeline to analyse time-course gene expression data

### 2019-05-15 manuscript.tex
Deploy NelleV/2019timecourse-rnaseq-pipeline to github.com/NelleV/2019timecourse-rnaseq-pipeline.git:gh-pages

\begin{itemize}
\tightlist
\item
  Add legends to the plots that don't have.
\item
  Currently have two functions ``plot\_genes'', ``plot\_centroids'', that do exactly
  the same thing. Find a better name for both of them
\item
  switch to camel case function names
\item
  shorten function names
\item
  Remove the namespace call
\item
  Move the normalization to a supplementary data
\item
  Make clear that it applies to RNASeq and how to call the different functions
  toRNASeq
\item
  Check that clustering can deal with count data.
\end{itemize}

@@

\textbf{EAP: General comment. Your function names are incredibly long compared to
standard functions. Also, will want to get rid of \texttt{moanin::} in front of
functions. It's more precise, but an average user doesn't ever experience the
\texttt{::} command; usually just package creaters.}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{variance_cutoff =}\StringTok{ }\FloatTok{0.5}
\NormalTok{variance_per_genes =}\StringTok{ }\KeywordTok{apply}\NormalTok{(data, }\DecValTok{1}\NormalTok{, mad)}
\NormalTok{min_variance =}\StringTok{ }\NormalTok{stats}\OperatorTok{::}\KeywordTok{quantile}\NormalTok{(variance_per_genes, }\KeywordTok{c}\NormalTok{(variance_cutoff))}
\NormalTok{variance_filtered_data =}\StringTok{ }\NormalTok{data[variance_per_genes }\OperatorTok{>}\StringTok{ }\NormalTok{min_variance,]}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{pca_data =}\StringTok{ }\NormalTok{stats}\OperatorTok{::}\KeywordTok{prcomp}\NormalTok{(}\KeywordTok{t}\NormalTok{(variance_filtered_data), }\DataTypeTok{rank=}\DecValTok{3}\NormalTok{, }\DataTypeTok{center=}\OtherTok{TRUE}\NormalTok{, }\DataTypeTok{scale=}\OtherTok{TRUE}\NormalTok{) }
\NormalTok{percent_var =}\StringTok{ }\KeywordTok{round}\NormalTok{(}\DecValTok{100} \OperatorTok{*}\StringTok{ }\KeywordTok{attr}\NormalTok{(pca_data, }\StringTok{"percentVar"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{graphics}\OperatorTok{::}\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow=}\KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{mar=}\KeywordTok{c}\NormalTok{(}\FloatTok{2.5}\NormalTok{, }\FloatTok{2.5}\NormalTok{, }\FloatTok{2.5}\NormalTok{, }\FloatTok{2.5}\NormalTok{))}
\NormalTok{graphics}\OperatorTok{::}\KeywordTok{plot}\NormalTok{(}
\NormalTok{    pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC2"}\NormalTok{], pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC1"}\NormalTok{],}
    \DataTypeTok{col=}\NormalTok{ann_colors}\OperatorTok{$}\NormalTok{Group[meta}\OperatorTok{$}\NormalTok{Group],}
    \DataTypeTok{pch=}\NormalTok{ann_markers}\OperatorTok{$}\NormalTok{Replicate[}\KeywordTok{as.factor}\NormalTok{(meta}\OperatorTok{$}\NormalTok{Replicate)],}
    \DataTypeTok{xlab=}\StringTok{"PC2"}\NormalTok{, }\DataTypeTok{ylab=}\StringTok{"PC1"}\NormalTok{)}
\NormalTok{graphics}\OperatorTok{::}\KeywordTok{plot}\NormalTok{(}
\NormalTok{    pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC2"}\NormalTok{], pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC1"}\NormalTok{],}
    \DataTypeTok{col=}\NormalTok{ann_colors}\OperatorTok{$}\NormalTok{Timepoint[}\KeywordTok{as.factor}\NormalTok{(meta}\OperatorTok{$}\NormalTok{Timepoint)],}
    \DataTypeTok{pch=}\NormalTok{ann_markers}\OperatorTok{$}\NormalTok{Replicate[}\KeywordTok{as.factor}\NormalTok{(meta}\OperatorTok{$}\NormalTok{Replicate)],}
    \DataTypeTok{xlab=}\StringTok{"PC2"}\NormalTok{, }\DataTypeTok{ylab=}\StringTok{"PC1"}\NormalTok{)}

\NormalTok{graphics}\OperatorTok{::}\KeywordTok{plot}\NormalTok{(}
\NormalTok{    pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC2"}\NormalTok{], pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC3"}\NormalTok{],}
    \DataTypeTok{col=}\NormalTok{ann_colors}\OperatorTok{$}\NormalTok{Group[meta}\OperatorTok{$}\NormalTok{Group],}
    \DataTypeTok{pch=}\NormalTok{ann_markers}\OperatorTok{$}\NormalTok{Replicate[}\KeywordTok{as.factor}\NormalTok{(meta}\OperatorTok{$}\NormalTok{Replicate)],}
    \DataTypeTok{xlab=}\StringTok{"PC2"}\NormalTok{, }\DataTypeTok{ylab=}\StringTok{"PC3"}\NormalTok{)}
\NormalTok{graphics}\OperatorTok{::}\KeywordTok{plot}\NormalTok{(}
\NormalTok{    pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC2"}\NormalTok{], pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC3"}\NormalTok{],}
    \DataTypeTok{col=}\NormalTok{ann_colors}\OperatorTok{$}\NormalTok{Timepoint[}\KeywordTok{as.factor}\NormalTok{(meta}\OperatorTok{$}\NormalTok{Timepoint)],}
    \DataTypeTok{pch=}\NormalTok{ann_markers}\OperatorTok{$}\NormalTok{Replicate[}\KeywordTok{as.factor}\NormalTok{(meta}\OperatorTok{$}\NormalTok{Replicate)],}
    \DataTypeTok{xlab=}\StringTok{"PC2"}\NormalTok{, }\DataTypeTok{ylab=}\StringTok{"PC3"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{variance_filtered_data =}\StringTok{ }\NormalTok{variance_filtered_data[, ord]}
\NormalTok{data_corr =}\StringTok{ }\NormalTok{stats}\OperatorTok{::}\KeywordTok{cor}\NormalTok{(variance_filtered_data, }\DataTypeTok{method=}\StringTok{"pearson"}\NormalTok{)}
\NormalTok{data_corr_meta =}\StringTok{ }\NormalTok{meta[ord, ]}
\NormalTok{data_corr_meta}\OperatorTok{$}\NormalTok{Timepoint =}\StringTok{ }\KeywordTok{as.factor}\NormalTok{(data_corr_meta}\OperatorTok{$}\NormalTok{Timepoint)}

\NormalTok{NMF}\OperatorTok{::}\KeywordTok{aheatmap}\NormalTok{(}
\NormalTok{    data_corr,}
    \DataTypeTok{Colv=}\OtherTok{NA}\NormalTok{, }\DataTypeTok{Rowv=}\OtherTok{NA}\NormalTok{,}
    \DataTypeTok{annCol=}\NormalTok{data_corr_meta[, }\KeywordTok{c}\NormalTok{(}\StringTok{"Group"}\NormalTok{, }\StringTok{"Timepoint"}\NormalTok{)],}
    \DataTypeTok{annRow=}\NormalTok{data_corr_meta[, }\KeywordTok{c}\NormalTok{(}\StringTok{"Group"}\NormalTok{, }\StringTok{"Timepoint"}\NormalTok{)],}
    \DataTypeTok{annLegend=}\OtherTok{TRUE}\NormalTok{, }
    \DataTypeTok{annColors=}\NormalTok{ann_colors,}
    \DataTypeTok{main=}\StringTok{"Correlation plot"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

If no formula is provided, it will default to the following:
\texttt{formula\ =\ \textasciitilde{}Group:splines::ns(Timepoint,\ df=degrees\_of\_fredoom)\ +\ Group\ +\ 0}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{splines_model =}\StringTok{ }\NormalTok{moanin}\OperatorTok{::}\KeywordTok{create_splines_model}\NormalTok{(meta, }\DataTypeTok{degrees_of_freedom=}\DecValTok{6}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Define contrast   }
\NormalTok{weekly_de_analysis =}\StringTok{  }\NormalTok{moanin}\OperatorTok{::}\KeywordTok{DE_timepoints}\NormalTok{(}
\NormalTok{     data, splines_model, contrasts,}
     \DataTypeTok{use_voom_weights=}\OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{contrasts =}\StringTok{ }\NormalTok{limma}\OperatorTok{::}\KeywordTok{makeContrasts}\NormalTok{(}
    \DataTypeTok{contrasts=}\NormalTok{timecourse_contrasts,}
    \DataTypeTok{levels=}\KeywordTok{levels}\NormalTok{(meta}\OperatorTok{$}\NormalTok{Group))}

\CommentTok{# The function takes the data (data.frame or named matrix), the meta data}
\CommentTok{# (data.frame containing a timepoint and group column, the first corresponding}
\CommentTok{#<c2><a0>to the time-course information, the latter corresponding to the}
\CommentTok{# treatment).}
\NormalTok{pvalues =}\StringTok{ }\NormalTok{moanin}\OperatorTok{::}\KeywordTok{DE_timecourse}\NormalTok{(}
\NormalTok{    data, splines_model, contrasts,}
    \DataTypeTok{use_voom_weights=}\OtherTok{FALSE}\NormalTok{)}
\NormalTok{qvalues =}\StringTok{ }\KeywordTok{apply}\NormalTok{(pvalues, }\DecValTok{2}\NormalTok{, stats}\OperatorTok{::}\NormalTok{p.adjust)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{log_fold_change_timepoints =}\StringTok{ }\NormalTok{moanin}\OperatorTok{::}\KeywordTok{estimate_log_fold_change}\NormalTok{(}
\NormalTok{    data, splines_model, contrasts,  }\DataTypeTok{method=}\StringTok{"timely"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{log_fold_change_epicon =}\StringTok{ }\NormalTok{moanin}\OperatorTok{::}\KeywordTok{estimate_log_fold_change}\NormalTok{(}
\NormalTok{    data, splines_model, contrasts,  }\DataTypeTok{method=}\StringTok{"epicon"}\NormalTok{)}

\NormalTok{log_fold_change_sum =}\StringTok{ }\NormalTok{moanin}\OperatorTok{::}\KeywordTok{estimate_log_fold_change}\NormalTok{(}
\NormalTok{    data, splines_model, contrasts,  }\DataTypeTok{method=}\StringTok{"sum"}\NormalTok{)}

\NormalTok{log_fold_change_max =}\StringTok{ }\NormalTok{moanin}\OperatorTok{::}\KeywordTok{estimate_log_fold_change}\NormalTok{(}
\NormalTok{    data, splines_model, contrasts, }\DataTypeTok{method=}\StringTok{"max"}\NormalTok{)}

\NormalTok{log_fold_change_min =}\StringTok{ }\NormalTok{moanin}\OperatorTok{::}\KeywordTok{estimate_log_fold_change}\NormalTok{(}
\NormalTok{    data, splines_model, contrasts, }\DataTypeTok{method=}\StringTok{"min"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{top_DE_genes_pval =}\StringTok{ }\KeywordTok{names}\NormalTok{(}\KeywordTok{sort}\NormalTok{(pvalue)[}\DecValTok{1}\OperatorTok{:}\DecValTok{12}\NormalTok{])}
\NormalTok{moanin}\OperatorTok{::}\KeywordTok{plot_genes}\NormalTok{(data[top_DE_genes_pval, ], splines_model,}
           \DataTypeTok{colors=}\NormalTok{ann_colors}\OperatorTok{$}\NormalTok{Group, }\DataTypeTok{smooth=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{top_DE_genes_lfc =}\StringTok{ }\KeywordTok{names}\NormalTok{(}
    \KeywordTok{sort}\NormalTok{(}\KeywordTok{abs}\NormalTok{(lfc_epicon),}
     \DataTypeTok{decreasing=}\OtherTok{TRUE}\NormalTok{)[}\DecValTok{1}\OperatorTok{:}\DecValTok{12}\NormalTok{])}
\NormalTok{moanin}\OperatorTok{::}\KeywordTok{plot_genes}\NormalTok{(data[top_DE_genes_lfc, ], splines_model,}
           \DataTypeTok{colors=}\NormalTok{ann_colors}\OperatorTok{$}\NormalTok{Group, }\DataTypeTok{smooth=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Then rank by fisher's p-value and take max the number of genes of interest}
\CommentTok{# Filter out q-values for the pvalues table}
\NormalTok{fishers_pval =}\StringTok{ }\NormalTok{moanin}\OperatorTok{::}\KeywordTok{pvalues_fisher_method}\NormalTok{(pvalues)}
\NormalTok{fishers_qval =}\StringTok{ }\NormalTok{stats}\OperatorTok{::}\KeywordTok{p.adjust}\NormalTok{(fishers_pval)}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# First fit the kmeans clusters}
\NormalTok{kmeans_clusters =}\StringTok{ }\NormalTok{moanin}\OperatorTok{::}\KeywordTok{splines_kmeans}\NormalTok{(}
\NormalTok{    y, splines_model, }\DataTypeTok{n_clusters=}\DecValTok{8}\NormalTok{,}
    \DataTypeTok{random_seed=}\DecValTok{42}\NormalTok{,}
    \DataTypeTok{n_init=}\DecValTok{20}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{moanin}\OperatorTok{::}\KeywordTok{plot_centroids}\NormalTok{(kmeans_clusters}\OperatorTok{$}\NormalTok{centroids, splines_model,}
               \DataTypeTok{colors=}\NormalTok{ann_colors}\OperatorTok{$}\NormalTok{Group,}
               \DataTypeTok{smooth=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{#<c2><a0>Then assign scores and labels to all the data, using a goodness-of-fit}
\CommentTok{# scoring function.}
\NormalTok{scores_and_labels =}\StringTok{ }\NormalTok{moanin}\OperatorTok{::}\KeywordTok{splines_kmeans_score_and_label}\NormalTok{(}
\NormalTok{    data, kmeans_clusters)}
\NormalTok{labels =}\StringTok{ }\NormalTok{scores_and_labels}\OperatorTok{$}\NormalTok{labels}

@@

\NormalTok{res =}\StringTok{ }\NormalTok{NMF}\OperatorTok{::}\KeywordTok{aheatmap}\NormalTok{(}
\NormalTok{    data_to_plot,}
    \DataTypeTok{Colv=}\OtherTok{NA}\NormalTok{,}
    \DataTypeTok{color=}\StringTok{"YlGnBu"}\NormalTok{,}
    \DataTypeTok{annCol=}\NormalTok{submeta[,}
                \KeywordTok{c}\NormalTok{(}\StringTok{"Group"}\NormalTok{, }\StringTok{"Timepoint"}\NormalTok{)],}
    \DataTypeTok{annLegend=}\OtherTok{FALSE}\NormalTok{,}
    \DataTypeTok{annColors=}\NormalTok{ann_colors,}
    \DataTypeTok{main=}\KeywordTok{paste}\NormalTok{(}\StringTok{"Cluster"}\NormalTok{, cluster_to_plot, }\StringTok{"(raw)"}\NormalTok{),}
    \DataTypeTok{treeheight=}\DecValTok{0}\NormalTok{, }\DataTypeTok{legend=}\OtherTok{FALSE}\NormalTok{)}

>>>>

\begin{itemize}
\tightlist
\item
  Add legends to the plots that don't have.
\item
  Currently have two functions ``plot\_genes'', ``plot\_centroids'', that do exactly
  the same thing. Find a better name for both of them
\item
  switch to camel case function names
\item
  Remove the namespace call
\item
  Make clear that it applies to RNASeq and how to call the different functions
  toRNASeq
\item
  Check that clustering can deal with count data.
\end{itemize}

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{variance_cutoff =}\StringTok{ }\FloatTok{0.5}
\NormalTok{variance_per_genes =}\StringTok{ }\KeywordTok{apply}\NormalTok{(data, }\DecValTok{1}\NormalTok{, mad)}
\NormalTok{min_variance =}\StringTok{ }\KeywordTok{quantile}\NormalTok{(variance_per_genes, }\KeywordTok{c}\NormalTok{(variance_cutoff))}
\NormalTok{variance_filtered_data =}\StringTok{ }\NormalTok{data[variance_per_genes }\OperatorTok{>}\StringTok{ }\NormalTok{min_variance,]}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{pca_data =}\StringTok{ }\KeywordTok{prcomp}\NormalTok{(}\KeywordTok{t}\NormalTok{(variance_filtered_data), }\DataTypeTok{rank=}\DecValTok{3}\NormalTok{, }\DataTypeTok{center=}\OtherTok{TRUE}\NormalTok{, }\DataTypeTok{scale=}\OtherTok{TRUE}\NormalTok{) }
\NormalTok{percent_var =}\StringTok{ }\KeywordTok{round}\NormalTok{(}\DecValTok{100} \OperatorTok{*}\StringTok{ }\KeywordTok{attr}\NormalTok{(pca_data, }\StringTok{"percentVar"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow=}\KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{mar=}\KeywordTok{c}\NormalTok{(}\FloatTok{2.5}\NormalTok{, }\FloatTok{2.5}\NormalTok{, }\FloatTok{2.5}\NormalTok{, }\FloatTok{2.5}\NormalTok{))}
\KeywordTok{plot}\NormalTok{(}
\NormalTok{    pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC2"}\NormalTok{], pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC1"}\NormalTok{],}
    \DataTypeTok{col=}\NormalTok{ann_colors}\OperatorTok{$}\NormalTok{Group[meta}\OperatorTok{$}\NormalTok{Group],}
    \DataTypeTok{pch=}\NormalTok{ann_markers}\OperatorTok{$}\NormalTok{Replicate[}\KeywordTok{as.factor}\NormalTok{(meta}\OperatorTok{$}\NormalTok{Replicate)],}
    \DataTypeTok{xlab=}\StringTok{"PC2"}\NormalTok{, }\DataTypeTok{ylab=}\StringTok{"PC1"}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(}
\NormalTok{    pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC2"}\NormalTok{], pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC1"}\NormalTok{],}
    \DataTypeTok{col=}\NormalTok{ann_colors}\OperatorTok{$}\NormalTok{Timepoint[}\KeywordTok{as.factor}\NormalTok{(meta}\OperatorTok{$}\NormalTok{Timepoint)],}
    \DataTypeTok{pch=}\NormalTok{ann_markers}\OperatorTok{$}\NormalTok{Replicate[}\KeywordTok{as.factor}\NormalTok{(meta}\OperatorTok{$}\NormalTok{Replicate)],}
    \DataTypeTok{xlab=}\StringTok{"PC2"}\NormalTok{, }\DataTypeTok{ylab=}\StringTok{"PC1"}\NormalTok{)}

\KeywordTok{plot}\NormalTok{(}
\NormalTok{    pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC2"}\NormalTok{], pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC3"}\NormalTok{],}
    \DataTypeTok{col=}\NormalTok{ann_colors}\OperatorTok{$}\NormalTok{Group[meta}\OperatorTok{$}\NormalTok{Group],}
    \DataTypeTok{pch=}\NormalTok{ann_markers}\OperatorTok{$}\NormalTok{Replicate[}\KeywordTok{as.factor}\NormalTok{(meta}\OperatorTok{$}\NormalTok{Replicate)],}
    \DataTypeTok{xlab=}\StringTok{"PC2"}\NormalTok{, }\DataTypeTok{ylab=}\StringTok{"PC3"}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(}
\NormalTok{    pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC2"}\NormalTok{], pca_data}\OperatorTok{$}\NormalTok{x[, }\StringTok{"PC3"}\NormalTok{],}
    \DataTypeTok{col=}\NormalTok{ann_colors}\OperatorTok{$}\NormalTok{Timepoint[}\KeywordTok{as.factor}\NormalTok{(meta}\OperatorTok{$}\NormalTok{Timepoint)],}
    \DataTypeTok{pch=}\NormalTok{ann_markers}\OperatorTok{$}\NormalTok{Replicate[}\KeywordTok{as.factor}\NormalTok{(meta}\OperatorTok{$}\NormalTok{Replicate)],}
    \DataTypeTok{xlab=}\StringTok{"PC2"}\NormalTok{, }\DataTypeTok{ylab=}\StringTok{"PC3"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{variance_filtered_data =}\StringTok{ }\NormalTok{variance_filtered_data[, ord]}
\NormalTok{data_corr =}\StringTok{ }\KeywordTok{cor}\NormalTok{(variance_filtered_data, }\DataTypeTok{method=}\StringTok{"pearson"}\NormalTok{)}
\NormalTok{data_corr_meta =}\StringTok{ }\NormalTok{meta[ord, ]}
\NormalTok{data_corr_meta}\OperatorTok{$}\NormalTok{Timepoint =}\StringTok{ }\KeywordTok{as.factor}\NormalTok{(data_corr_meta}\OperatorTok{$}\NormalTok{Timepoint)}

\KeywordTok{aheatmap}\NormalTok{(}
\NormalTok{    data_corr,}
    \DataTypeTok{Colv=}\OtherTok{NA}\NormalTok{, }\DataTypeTok{Rowv=}\OtherTok{NA}\NormalTok{,}
    \DataTypeTok{annCol=}\NormalTok{data_corr_meta[, }\KeywordTok{c}\NormalTok{(}\StringTok{"Group"}\NormalTok{, }\StringTok{"Timepoint"}\NormalTok{)],}
    \DataTypeTok{annRow=}\NormalTok{data_corr_meta[, }\KeywordTok{c}\NormalTok{(}\StringTok{"Group"}\NormalTok{, }\StringTok{"Timepoint"}\NormalTok{)],}
    \DataTypeTok{annLegend=}\OtherTok{TRUE}\NormalTok{, }
    \DataTypeTok{annColors=}\NormalTok{ann_colors,}
    \DataTypeTok{main=}\StringTok{"Correlation plot"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

If no formula is provided, it will default to the following:
\texttt{formula\ =\ \textasciitilde{}Group:ns(Timepoint,\ df=degrees\_of\_fredoom)\ +\ Group\ +\ 0}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{splines_model =}\StringTok{ }\KeywordTok{create_splines_model}\NormalTok{(meta, }\DataTypeTok{degrees_of_freedom=}\DecValTok{6}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Define contrast   }
\NormalTok{weekly_de_analysis =}\StringTok{ }\KeywordTok{DE_timepoints}\NormalTok{(}
\NormalTok{     data, splines_model, contrasts,}
     \DataTypeTok{use_voom_weights=}\OtherTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{contrasts =}\StringTok{ }\KeywordTok{makeContrasts}\NormalTok{(}
    \DataTypeTok{contrasts=}\NormalTok{timecourse_contrasts,}
    \DataTypeTok{levels=}\KeywordTok{levels}\NormalTok{(meta}\OperatorTok{$}\NormalTok{Group))}

\CommentTok{# The function takes the data (data.frame or named matrix), the meta data}
\CommentTok{# (data.frame containing a timepoint and group column, the first corresponding}
\CommentTok{#<c2><a0>to the time-course information, the latter corresponding to the}
\CommentTok{# treatment).}
\NormalTok{pvalues =}\StringTok{ }\KeywordTok{DE_timecourse}\NormalTok{(}
\NormalTok{    data, splines_model, contrasts,}
    \DataTypeTok{use_voom_weights=}\OtherTok{FALSE}\NormalTok{)}
\NormalTok{qvalues =}\StringTok{ }\KeywordTok{apply}\NormalTok{(pvalues, }\DecValTok{2}\NormalTok{, p.adjust)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{log_fold_change_timepoints =}\StringTok{ }\KeywordTok{estimate_log_fold_change}\NormalTok{(}
\NormalTok{    data, splines_model, contrasts,  }\DataTypeTok{method=}\StringTok{"timely"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{log_fold_change_epicon =}\StringTok{ }\KeywordTok{estimate_log_fold_change}\NormalTok{(}
\NormalTok{    data, splines_model, contrasts,  }\DataTypeTok{method=}\StringTok{"epicon"}\NormalTok{)}

\NormalTok{log_fold_change_sum =}\StringTok{ }\KeywordTok{estimate_log_fold_change}\NormalTok{(}
\NormalTok{    data, splines_model, contrasts,  }\DataTypeTok{method=}\StringTok{"sum"}\NormalTok{)}

\NormalTok{log_fold_change_max =}\StringTok{ }\KeywordTok{estimate_log_fold_change}\NormalTok{(}
\NormalTok{    data, splines_model, contrasts, }\DataTypeTok{method=}\StringTok{"max"}\NormalTok{)}

\NormalTok{log_fold_change_min =}\StringTok{ }\KeywordTok{estimate_log_fold_change}\NormalTok{(}
\NormalTok{    data, splines_model, contrasts, }\DataTypeTok{method=}\StringTok{"min"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{top_DE_genes_pval =}\StringTok{ }\KeywordTok{names}\NormalTok{(}\KeywordTok{sort}\NormalTok{(pvalue)[}\DecValTok{1}\OperatorTok{:}\DecValTok{12}\NormalTok{])}
\KeywordTok{plot_genes}\NormalTok{(data[top_DE_genes_pval, ], splines_model,}
           \DataTypeTok{colors=}\NormalTok{ann_colors}\OperatorTok{$}\NormalTok{Group, }\DataTypeTok{smooth=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{top_DE_genes_lfc =}\StringTok{ }\KeywordTok{names}\NormalTok{(}
    \KeywordTok{sort}\NormalTok{(}\KeywordTok{abs}\NormalTok{(lfc_epicon),}
     \DataTypeTok{decreasing=}\OtherTok{TRUE}\NormalTok{)[}\DecValTok{1}\OperatorTok{:}\DecValTok{12}\NormalTok{])}
\KeywordTok{plot_genes}\NormalTok{(data[top_DE_genes_lfc, ], splines_model,}
           \DataTypeTok{colors=}\NormalTok{ann_colors}\OperatorTok{$}\NormalTok{Group, }\DataTypeTok{smooth=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Then rank by fisher's p-value and take max the number of genes of interest}
\CommentTok{# Filter out q-values for the pvalues table}
\NormalTok{fishers_pval =}\StringTok{ }\KeywordTok{pvalues_fisher_method}\NormalTok{(pvalues)}
\NormalTok{fishers_qval =}\StringTok{ }\KeywordTok{p.adjust}\NormalTok{(fishers_pval)}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# First fit the kmeans clusters}
\NormalTok{kmeans_clusters =}\StringTok{ }\KeywordTok{splines_kmeans}\NormalTok{(}
\NormalTok{    y, splines_model, }\DataTypeTok{n_clusters=}\DecValTok{8}\NormalTok{,}
    \DataTypeTok{random_seed=}\DecValTok{42}\NormalTok{,}
    \DataTypeTok{n_init=}\DecValTok{20}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot_centroids}\NormalTok{(kmeans_clusters}\OperatorTok{$}\NormalTok{centroids, splines_model,}
               \DataTypeTok{colors=}\NormalTok{ann_colors}\OperatorTok{$}\NormalTok{Group,}
               \DataTypeTok{smooth=}\OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{#<c2><a0>Then assign scores and labels to all the data, using a goodness-of-fit}
\CommentTok{# scoring function.}
\NormalTok{scores_and_labels =}\StringTok{ }\KeywordTok{splines_kmeans_score_and_label}\NormalTok{(}
\NormalTok{    data, kmeans_clusters)}
\NormalTok{labels =}\StringTok{ }\NormalTok{scores_and_labels}\OperatorTok{$}\NormalTok{labels}

@@

\NormalTok{res =}\StringTok{ }\KeywordTok{aheatmap}\NormalTok{(}
\NormalTok{    data_to_plot,}
    \DataTypeTok{Colv=}\OtherTok{NA}\NormalTok{,}
    \DataTypeTok{color=}\StringTok{"YlGnBu"}\NormalTok{,}
    \DataTypeTok{annCol=}\NormalTok{submeta[,}
                \KeywordTok{c}\NormalTok{(}\StringTok{"Group"}\NormalTok{, }\StringTok{"Timepoint"}\NormalTok{)],}
    \DataTypeTok{annLegend=}\OtherTok{FALSE}\NormalTok{,}
    \DataTypeTok{annColors=}\NormalTok{ann_colors,}
    \DataTypeTok{main=}\KeywordTok{paste}\NormalTok{(}\StringTok{"Cluster"}\NormalTok{, cluster_to_plot, }\StringTok{"(raw)"}\NormalTok{),}
    \DataTypeTok{treeheight=}\DecValTok{0}\NormalTok{, }\DataTypeTok{legend=}\OtherTok{FALSE}\NormalTok{)}

\CommentTok{# Now use the results of the previous call to aheatmap to reorder the genes.}
\KeywordTok{aheatmap}\NormalTok{(}
\NormalTok{    moanin}\OperatorTok{:::}\KeywordTok{rescale_values}\NormalTok{(data_to_plot)[res}\OperatorTok{$}\NormalTok{rowInd,],}
    \DataTypeTok{Colv=}\OtherTok{NA}\NormalTok{,}
    \DataTypeTok{Rowv=}\OtherTok{NA}\NormalTok{,}
    \DataTypeTok{annCol=}\NormalTok{submeta[,}
                \KeywordTok{c}\NormalTok{(}\StringTok{"Group"}\NormalTok{, }\StringTok{"Timepoint"}\NormalTok{)],}
    \DataTypeTok{annLegend=}\OtherTok{TRUE}\NormalTok{,}
    \DataTypeTok{annColors=}\NormalTok{ann_colors,}
    \DataTypeTok{main=}\KeywordTok{paste}\NormalTok{(}\StringTok{"Cluster"}\NormalTok{, cluster_to_plot, }\StringTok{"(rescaled)"}\NormalTok{),}
    \DataTypeTok{treeheight=}\DecValTok{0}\NormalTok{)}
\end{Highlighting}
\end{Shaded}
